Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-05-03 GWLDCII.tex
fixed 3.4, addressed most of the notes in Susama's file

A Wilson loop diagram is depicted as a circle, with the vertices $V$ equally spaced along the edge of the circle and listed counterclockwise in their cyclic order. The arc between two consecutive vertices is referred to as an {\bf edge} of the diagram, and the $i$th edge of $W$ is the edge from vertex $i$ to its successor. Each propagator $(i,j) \in \cP$ is denoted by a wavy line inside the circle joining edge $i$ and edge $j$.

@@

\begin{dfn} \label{subdiagramdfn}
Let $W = (\cP, V)$ be an admissible Wilson loop diagram. Then $W' = (Q,U)$ is a {\bf subdiagram} of $W$ if $Q \subseteq \cP$ and $V(Q) \subseteq U \subseteq V$.
\end{dfn}

@@

These conditions are imposed by the physical interpretation of Wilson loop diagrams (see for example \hlfix{[ref]}{probably for Susama: is there a sensible paper to cite here?}). We therefore restrict our attention in this paper to admissible Wilson loop diagrams and their subdiagrams only.

Note that a subdiagram $(Q,U)$ of an admissible Wilson loop diagram need not be admissible itself: it inherits conditions 2 and 3 automatically, but we could have $|U| = |Q| + 3$. We call diagrams of this type {\bf weakly admissible}; in the context of this paper, they occur only when we take an admissible diagram and delete one or more non-supporting vertices.

@@

The Wilson loop diagram also defines a volume form on $\Sigma(W)$, namely 
\[ \Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V_{p_r}} \textrm{d}c_{p_r}}{R(W)} \;. \]
The interested reader is referred to \hlfix{[ref]}{any paper that talks about the volume form} for more information about the volume form. Of interest to us in this paper is the denominator $R(W)$, which we define next.

\begin{dfn}\label{def R(W)}
Let $W = (\cP,V)$ be an admissible Wilson loop diagram, and fix an edge $e$. Let $\{p_1 \ldots p_r\}$ be the propagators with one end lying on edge $e$, ordered so that $p_1$ is closest to the vertex $e$, $p_r$ closest to $e+1$, and $p_i$ is closer to $e$ than $p_{i+1}$ is. Define 
\[ R_e =  c_{p_1,e+1} \prod_{j= 1}^{r-1} \left((c_{p_j,e} c_{p_{j+1},e+1} - c_{p_{j+1},e} c_{p_{j},e+1} ) \right) c_{p_r,e}\;\]
to be the component of $R(W)$ associated to edge $e$. Note that if $r = 1$ (i.e. there is exactly one propagator lying on edge $e$) then this expression simplifies to $R_e = c_{p,e} c_{p,e+1}$. The denominator $R(W)$ is defined to be the product of all of the $R_e$, i.e. 
\[R(W) = \prod_{e \in V} R_e.\]
\end{dfn}
Understanding the polynomial $R(W)$ (and hence the singularities of the volume form $\Omega(W)$) \hlfix{is important for the physics}{help?}. In Section~\ref{sec poles} we give a simple algorithm to read $R(W)$ directly from the Wilson loop diagram.

@@

For each $j \in [n]$, we can define a total order $\leq_j$ on the interval $[n]$ by
\[ j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\]
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the {\bf Gale order} \cite{Gale}), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we define
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

@@

By studying $M(W)$ in terms of the behavior of the propagators in $W$, Agarwala and Amat also showed that $M(W)$ is in fact a positroid whenever $W$ is admissible \cite[Corollary 3.39]{wilsonloop}. In theory, this allows us to leverage geometric and combinatorial results about $\Gr(k,n)$ to study Wilson loop diagrams.

@@

@@

In this section we will give an algorithm to go directly from a Wilson loop diagram to the Grassmann necklace of its positroid. This is a useful result in and of itself, as it greatly simplifies the process of identifying the positroid associated to a given WLD. It also allows us to examine how the positroid changes as we make minor changes to the Wilson loop diagram; indeed, the techniques we develop in this section to verify that our algorithm produces the correct Grassmann necklace will also be used repeatedly in later sections.

@@

@@

@@

\begin{algorithm}\label{alg:put GN on WLD}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. This gives an algorithm for constructing the set $I_j$, for $j \in [n]$.

\begin{enumerate}
\item Fix a vertex $j \in [n]$. Set $j:=i$ and $I_i = \emptyset$.
\item While $\cP \neq \emptyset$, perform the following steps:
\begin{enumerate}
\item \textbf{Step $j$ for vertex $i$}: If $\Prop(j) = \emptyset$, do nothing. Else, let $p \in \Prop(j)$ be the clockwise most propagator supported on $j$. Write $I_i = I_i\cup \{j\}$, and delete propagator $p$ from the diagram.
\item Increment $j$ by 1 and repeat from (a).
\end{enumerate}
\end{enumerate}
\end{algorithm}

@@

To construct $I_1$ we start at vertex 1, so we begin with $i=1$, $m = 1$, and $I_1 = \emptyset$. 
\begin{itemize}
\item $j = 1$: Since $\Prop(1) = \{q,r\}$ and $r$ is the clockwise-most of these propagators, we let $I_1 = \{1\}$ and remove propagator $r$ from the diagram. 
\item $j = 2$: $\Prop(2) = \{p,q\}$, with $q$ being the clockwise-most of these propagators. So $I_1 = \{1,2\}$ and propagator $q$ is removed.
\item $j = 3$: $\Prop(3) = \{p\}$, so now $I_1 = \{1,2,3\}$ and propagator $p$ is removed. The only remaining propagator in the diagram is $s$.
\item $j = 4$: Since $p$ and $q$ were removed in earlier steps, we now have $\Prop(4) = \emptyset$.
\item $j = 5$: $\Prop(5) = \{s\}$, and we have $I_1 = \{1,2,3,5\}$.
\end{itemize}
There are no propagators left in the diagram, so the algorithm terminates and we have ${I_1 = \{1,2,3,5\}}$, or $1235$ for short. Viewed as a function, we have
\[I_1: \cP \longrightarrow [n] : \quad I_1(p) = 3,\quad I_1(q) = 2, \quad I_1(r) = 1, \quad I_1(s) = 5.\]
Applying the algorithm for all 8 starting vertices, we obtain the sets
\[1235, 2356, 3456, 4567, 5671, 6712, 7812, 8123.\]
The reader can easily verify that this sequence of $k$-sets satisfies the definition of a Grassmann necklace, and can (with significantly more work) also verify that this Grassmann necklace defines the positroid $M(W)$ associated to the Wilson loop diagram $W$ above.
\end{eg}

We first verify that this algorithm always terminates in fewer than $n$ steps.

\begin{lem}\label{lem no fourth vertex}
Let $W$ be an admissible Wilson loop diagram containing at least one propagator. For any $i \in [n]$ and for any $p=(a,b)$ with $i\leq_i a <_i b$, we have $I_i(p) \neq b+1$.
\end{lem}
\begin{proof}
Suppose for contradiction that we have $p = (a,b)$ with $i\leq_i a <_i b$ and $I_{i}(p) = b+1$. We may choose $p$ such that $|[a+1,b]|$ is minimal amongst propagators with this property.

Since $I_i(p) \neq b$, there must exist a propagator $q$ inside of $(p,<_i)$ with $I_i(q) = b$. The propagator $q$ cannot end on the edge $(b-1,b)$, as this would contradict the minimality of $p$, so we have $q = (c,c+1,b,b+1)$ with $a <_i c <_i b$, and $I_i(q) = b$.

In order for $q$ to remain unassigned until vertex $b$, there must be another propagator $r$ with an end on $(c,c+1)$ and $I_i(r) = c+1$; the only way this can occur is if $r$ is outside $q$ but inside $p$. Now $r$ contributes its fourth vertex to $I_i$, again contradicting the minimality of $p$.
\end{proof}

\begin{cor}\label{GN alg well defined}
If $W$ is an admissible Wilson loop diagram with $k$ propagators, then Algorithm~\ref{alg:put GN on WLD} assigns exactly $k$ vertices to each $I_i$.
\end{cor}
\begin{proof}
It follows from the proof of Lemma~\ref{lem no fourth vertex} that Algorithm~\ref{alg:put GN on WLD} can never reach the fourth vertex of a propagator's support (with respect to the starting vertex). Therefore if the algorithm starts at vertex $i$, it must have assigned vertices to all propagators by the time it reaches $i-1$, ensuring that $I_i$ contains exactly $k$ distinct vertices.
\end{proof}

@@

Specifically, for each $i$ in the support of $p$, define the set
\[J_p^{(W)}(i) = \{m \in [n] \ | \ I^{(W)}_m(p) = i \},\]
the set of starting vertices $m$ such that $I_m^{(W)}$ assigns $p$ to $i$.

@@

Let $p=(i,j)$ be a propagator in an admissible WLD $W$ on $n$ vertices.  Then $J_p^{(W)}(i)$, $J_p^{(W)}(i+1)$, $J_p^{(W)}(j)$, and $J_p^{(W)}(j+1)$ are each non-empty cyclic intervals which partition $[n]$ and occur in the given cyclic order.
\end{lem}

@@

Consider the fourth case of Figure~\ref{fig 3 cases} first, as it is the easiest.  In this case, the vertices $\{l,l+1,l+2,l+3 \}$ which support $p$ in $W$ are non-supporting in $V$, so for every propagator $r$ in $V$ (including $q$) we have $J_r^{(V)} = J_r^{(W)}$; these are non-empty cyclic intervals in the correct order by the induction hypothesis.  Additionally, it is clear from Figure~\ref{fig 3 cases} that $J_p^{(W)}(l+a) = \{l+a\}$ for $a\in\{1,2,3\}$ and $J_p^{(W)}(l) = [n] \backslash \{l+1, l+2, l+3\}$. The result therefore holds in this case.

Now we proceed to consider the first three cases of Figure~\ref{fig 3 cases}.
We first describe $J_q^{(V)}(*)$, which can be handled identically for all three cases.

In $V$ there are no propagators inside $q$, so we see from Figure~\ref{fig V diagram} that
\[l, l+1, \ldots ,j \in J^{(V)}_q(j) \text{ (if $l$ exists) and }  j+1 \in J^{(V)}_q(j+1).\]
Note that $j+2 \not\in J^{(V)}_q(j+1)$ by Lemma~\ref{lem no fourth vertex}, so by the induction hypothesis we must have $J^{(V)}_q(j+1) = \{j+1\}$ and $j+2 \in J^{(V)}_q(i)$. Thus there exist vertices $d,e \in [j+2,i+1]$ with $d <e$, such that
\[J^{(V)}_q(i) = [j+2,d-1], \quad J^{(V)}_q(i+1) = [d,e-1], \quad J^{(V)}_q(j) = [e,j], \quad J^{(V)}_q(j+1) = \{j+1\},\]
and all intervals are non-empty.

We now consider what happens as we move from $V$ to each of the three remaining cases for $W$. We need to consider both $J^{(W)}_p$ and $J^{(W)}_r$ for $r \neq p$, since the addition of $p$ can have a knock-on effect on later steps in the algorithm.

@@

We first calculate $I^{(W)}_w$ for a starting vertex ${w \in [n]\backslash \{l, l+1, \ldots,j,j+1\}}$. Note that $p$ has no effect on other propagators for starting vertices in this range, while the value of $I^{(W)}_w(p)$ depends on how soon $q$ is assigned to a vertex, i.e. on the value of $I^{(V)}_w(q)$. Thus, if $w\in J_q^{(V)}(i)$ then
    \[
    I_w^{(W)}(r) =  \begin{cases}
        \max\{i+1, m\} & \text{if } r=p \\
        I_{w}^{(V)}(r) & \text{if } r\neq p
      \end{cases} 
    \]
    while if $w\in J_q^{(V)}(i+1)$ or $w\in J_q^{(V)}(j)$ then
    \[
    I_w^{(W)}(r) =  \begin{cases}
        l & \text{if } r=p \\
        I_{w}^{(V)}(r) & \text{if } r\neq p
      \end{cases} 
      \]
We also need to understand $I^{(W)}_w$ for $w \in \{l,l+1,\ldots,j,j+1\}$. For the majority of these vertices, we use the following observation: if $p$ is the first propagator to be assigned a value by $I_w^{(W)}$, then the remainder of $I_w^{(W)}$ proceeds identically to the assignments of $I^{(V)}_{w+1}$.  Thus we have for any $0\leq b <a$
    \[
    I_{l+b}^{(W)}(r) = \begin{cases}
      l+b & \text{if } r=p\\
      I_{l+b+1}^{(V)}(r) & \text{if } r\neq p
    \end{cases}
    \]
Similarly, if $j$ is in the support of $p$, then we have
     \[
       I_j^{(W)}(r) = \begin{cases}
         j & \text{if $r=p$ and $j$ is in the support of $p$}\\
         I_{j+1}^{(V)}(r) & \text{if $r\neq p$ and $j$ is in the support of $p$}\\
       \end{cases}
       \]
If $j$ is not in the support of $p$, then we must be in the second case of Figure~\ref{fig 3 cases} with two vertices in the right hand region.  In this case, if we start the algorithm at $j$ we need to know whether there will be any unassigned propagators other than $p$ when we reach vertex $i$, so as to know what $p$ contributes.

Consider the WLD $X$ formed from $V$ by moving the second end of $q$ to the edge $j-1$ instead of $j$.  $X$ is still admissible since we have not decreased the support of any set of propagators, so the induction hypothesis applies to it as well.  Note that $J_{j}^{(V)}(r) = J_{j}^{(X)}(r)$ for all $r\neq p$ and $J_{j}^{(X)}(r) = J_{j+1}^{(X)}(r)$ for $r\neq q,p$.  Additionally $I_{j+1}^{(X)}(q) = i$ by the induction hypothesis applied to $X$, and so if we start at $j+1$ and assign propagators to vertices according to the algorithm, when we reach vertex $i$ in $X$ all propagators other than $q$ must have been assigned.  Therefore if we start at $j$ in $W$, we first assign $q$ to $j$ then proceed to assign as in $X$ starting at $j+1$, and hence when we get to $i$ the only remaining unassigned propagator is $p$. Therefore
       \[
       I_j^{(W)}(r) = \begin{cases}

       m & \text{if $r =p$ and $j$ is not in the support of $p$} \\
       I_{j}^{(V)}(r) & \text{if $r\neq p$ and $j$ is not in the support of $p$}\end{cases}
       \] 
Finally, we consider what happens when we start the algorithm at vertex $j+1$. If $j+1$ is in the support of $p$ then we can argue as above to get
\[
       I_{j+1}^{(W)}(r)  = \begin{cases}
         j+1 & \text{if $r=p$ and $j+1$ is in the support of $p$}\\
         I_{j+2}^{(V)}(r) & \text{if $r\neq p$ and $j+1$ is in the support of $p$}.
       \end{cases}
       \]
Now suppose $j+1$ is not in the support of $p$. If we start at $j+1$ we need to know whether there are any unassigned propagators supported on edge $i$ when we reach vertex $i$. We already know that $J_q^{(V)}(j+1) = \{j+1\}$; in particular this means that $q$ contributes $i$ in $I_{j+2}^{(V)}$.  However the construction of $I^{(V)}_{j+1}$ first associates $q$ to $j+1$ and then proceeds identically to $I^{(V)}_{j+2}$.  In particular if $i$ was assigned in $I^{(V)}_{j+1}$, then it would not be available to assign to $q$ in $I^{(V)}_{j+2}$ as all other propagators supported at $i$ in $V$ come before $q$.

>>>>

A Wilson loop diagram is depicted as a circle, with the vertices $V$ arranged along the edge of the circle and listed counterclockwise in their cyclic order. The arc between two consecutive vertices is referred to as an {\bf edge} of the diagram, and the $i$th edge of $W$ is the edge from vertex $i$ to its successor. Each propagator $(i,j) \in \cP$ is denoted by a wavy line inside the circle joining edge $i$ and edge $j$.

@@

In particular, we will see in later sections that all four vertices in the support of a propagator are important in understanding the contribution of that propagator to the overall diagram. This motivates the following definition of a subdiagram, which can consist of any subset $Q$ of the propagators and any subset $U$ of the vertices of $W$ {\em as long as we retain all vertices in the support of $Q$}.

\begin{dfn} \label{subdiagramdfn}
Let $W = (\cP, V)$ be a Wilson loop diagram. Then $W' = (Q,U)$ is a {\bf subdiagram} of $W$ if $Q \subseteq \cP$ and $V(Q) \subseteq U \subseteq V$.
\end{dfn}

@@

These conditions are imposed by the physical interpretation of Wilson loop diagrams (see for example \cite{Adamo:2011pv,Adamo:2012xe,wilsonloop,LipsteinMason}). We therefore restrict our attention in this paper to admissible Wilson loop diagrams and their subdiagrams only.

Note that a subdiagram $(Q,U)$ of an admissible Wilson loop diagram need not be admissible itself: it inherits conditions 2 and 3 automatically, but we could have $|U| = |Q| + 3$. If a Wilson loop diagrams satisfies conditions 2 and 3 of Definition~\ref{admisdfn}, we call it {\bf weakly admissible}.

@@

The Wilson loop diagram also defines a volume form on $\Sigma(W)$, namely 
\[ \Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V_{p_r}} \textrm{d}c_{p_r}}{R(W)} \;. \]
The interested reader is referred to \cite{Adamo:2012xe,HeslopStewart,LipsteinMason} for more information about the volume form. Of interest to us in this paper is the denominator $R(W)$, which we define next.

\begin{dfn}\label{def R(W)}
Let $W = (\cP,V)$ be an admissible Wilson loop diagram, and fix an edge $e$. Let $\{p_1 \ldots p_r\}$ be the propagators with one end lying on edge $e$, ordered so that $p_1$ is closest to the vertex $e$, $p_r$ closest to $e+1$, and $p_i$ is closer to $e$ than $p_{i+1}$ is. Define 
\[ R_e =  c_{p_1,e+1} \prod_{j= 1}^{r-1} \left((c_{p_j,e} c_{p_{j+1},e+1} - c_{p_{j+1},e} c_{p_{j},e+1} ) \right) c_{p_r,e}\;\]
to be the component of $R(W)$ associated to edge $e$. Note that if $r = 1$ (i.e. there is exactly one propagator lying on edge $e$) then this expression simplifies to $R_e = c_{p,e} c_{p,e+1}$. The denominator $R(W)$ is defined to be the product of all of the $R_e$, i.e. 
\[R(W) = \prod_{e \in V} R_e.\]
\end{dfn}
We expect the volume form associated to the tree amplitudes to be finite \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}. Therefore, we expect that the singularities of $\Omega(W)$ will cancel on the boundaries of the simpleses the diagrams define, analogous to the analysis in \cite{Arkani-Hamed:2013jha}. While there has been work in this area in special cases \cite{Amplituhedronsquared,HeslopStewart,casestudy} there has not been a systematic study of the pole structure of these integrals.

In Section~\ref{sec poles} we give a simple algorithm to read $R(W)$ directly from the Wilson loop diagram, paving the way for a general study of these poles in future.

@@

For each $j \in [n]$, we can define a total order $<_j$ on the interval $[n]$ by
\[ j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\]
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the {\bf Gale order} \cite{Gale}), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we define
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

@@

By relating the behavior of the propagators in $W$ to the matroidal properties of $M(W)$, Agarwala and Amat also showed that $M(W)$ is in fact a positroid whenever $W$ is admissible \cite[Corollary 3.39]{wilsonloop}.

@@

\begin{rmk}
While this paper primarily approaches proofs from a combinatorial point of view, readers with a background in matroid theory are encouraged to keep the matroidal viewpoint in mind throughout; we expect that the two approaches will be complementary and will build upon each other. See [ref][paper 1] for some related results concerning the positroids of Wilson loop diagrams obtained by viewing them primarily as matroids.
\end{rmk}

@@

In this section we will give an algorithm (Algorithm~\ref{alg:put GN on WLD}) to go directly from a Wilson loop diagram to the Grassmann necklace of its positroid. This is a useful result in and of itself, as it greatly simplifies the process of identifying the positroid associated to a given WLD. It also allows us to examine how the positroid changes as we make minor changes to the Wilson loop diagram; indeed, the techniques we develop in this section to verify that our algorithm produces the correct Grassmann necklace will also be used repeatedly in later sections.

@@

\note{Sian thinks that a section roadmap here would be overkill: we gave a roadmap in section 1, each subsection has its own little intro roadmap, and I've added a ref to the algorithm above so that people can skip straight to it if they want}
%In Section~\ref{sec:propagator configs} we examine the structure of admissible Wilson loop diagrams and prove several technical lemmas about configurations of propagators appearing in these diagrams. In Section~\ref{sec:GN alg} we state the algorithm that extracts the Grassmann necklace of a given diagram (see Algorithm~\ref{alg:put GN on WLD}) and

@@

\hlfix{[picture]}{is it worth including a picture here?}

@@

\begin{algorithm}\label{alg:put GN on WLD}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. This gives an algorithm for constructing the set $I_i$, for $i \in [n]$.

\begin{enumerate}
\item Fix a vertex $i \in [n]$, the {\bf starting vertex}. Set $j:=i$ and $I_i = \emptyset$.
\item While $\cP \neq \emptyset$, perform the following steps:
\begin{enumerate}
\item {\em Step $j$ for starting vertex $i$}: If $\Prop(j) = \emptyset$, do nothing. Else, let $p \in \Prop(j)$ be the clockwise most propagator supported on $j$. Write $I_i = I_i\cup \{j\}$, and delete propagator $p$ from the diagram.
\item Increment $j$ by 1 and repeat from (a).
\end{enumerate}
\end{enumerate}
\end{algorithm}

@@

To construct $I_1$ we start at vertex 1, so we set $i=1$, $j = 1$, and $I_1 = \emptyset$. 
\begin{itemize}
\item $j = 1$: Since $\Prop(1) = \{q,r\}$ and $r$ is the clockwise-most of these propagators, we let $I_1 = \{1\}$ and remove propagator $r$ from the diagram. 
\item $j = 2$: $\Prop(2) = \{p,q\}$, with $q$ being the clockwise-most of these propagators. So $I_1 = \{1,2\}$ and propagator $q$ is removed.
\item $j = 3$: $\Prop(3) = \{p\}$, so now $I_1 = \{1,2,3\}$ and propagator $p$ is removed. The only remaining propagator in the diagram is $s$.
\item $j = 4$: Since $p$ and $q$ were removed in earlier steps, we now have $\Prop(4) = \emptyset$.
\item $j = 5$: $\Prop(5) = \{s\}$, and we have $I_1 = \{1,2,3,5\}$.
\end{itemize}
There are no propagators left in the diagram, so the algorithm terminates and we have ${I_1 = \{1,2,3,5\}}$, or $1235$ for short. Viewed as a function, we have
\[I_1: \cP \longrightarrow [n] : \quad I_1(p) = 3,\quad I_1(q) = 2, \quad I_1(r) = 1, \quad I_1(s) = 5.\]
Applying the algorithm for all 8 starting vertices, we obtain the sets
\[1235, 2356, 3456, 4567, 5671, 6712, 7812, 8123.\]
The reader can easily verify that this sequence of $k$-sets satisfies the definition of a Grassmann necklace, and can (with significantly more work) also verify that this Grassmann necklace defines the positroid $M(W)$ associated to the Wilson loop diagram $W$ above.
\end{eg}

We first show that the algorithm can never reach the final vertex in a propagator's support without having already assigned that propagator; in particular, this guarantees that the construction of $I_i$ always terminates in fewer than $n$ steps.

\begin{lem}\label{lem no fourth vertex}
Let $W$ be an admissible Wilson loop diagram containing at least one propagator. For any $i \in [n]$ and for any $p \in \cP$, $I_i(p)$ is not maximal (with respect to $<_i$) amongst the vertices that support $p$.
\end{lem}
\begin{proof}
Suppose for contradiction that we have $i \in [n]$ and $p \in \cP$ such that $I_i(p)$ is $<_i$-maximal in $V(p)$. There are two possibilities: either $p = (i-1,b)$ for some $b$ (so $i$ and $I_i(p) = i-1$ are the two ends of a single edge), or $p = (a,b)$ with $i\leq_i a <_i b$ and $I_i(p) = b+1$.

\includegraphics[width=15cm]{4th_vertex}

In the first case, observe that we must have another propagator $q = (c,b)$ with $c<_i b$ and $I_i(q) = b+1$, because otherwise $p$ would not survive until vertex $i-1$. Now $q$ also contributes its 4th vertex to $I_i$ and is an instance of the second case listed above, so it suffices to show that no admissible Wilson loop diagram admits this type of configuration.

So let $p = (a,b)$ with $a<_ib$ and $I_i(p) = b+1$, and choose $p$ such that $|[a+1,b]|$ is minimal amongst propagators of $W$ with this property.

Since $I_i(p) \neq b$, there must exist a propagator $q$ inside of $(p,<_i)$ with $I_i(q) = b$. The propagator $q$ cannot end on the edge $b-1$, as this would contradict the minimality of $p$, so we have $q = (d,b)$ with $a <_i d <_i b$, and $I_i(q) = b$.

In order for $q$ to remain unassigned until vertex $b$, there must be another propagator $r$ with an end on edge $d$ and $I_i(r) = d+1$; the only way this can occur is if $r$ is outside of $(q,<_i)$ but inside $(p,<_i)$. But now $r$ contributes its fourth vertex to $I_i$, again contradicting the minimality of $p$.
\end{proof}

\begin{cor}\label{GN alg well defined}
If $W$ is an admissible Wilson loop diagram with $k$ propagators, then Algorithm~\ref{alg:put GN on WLD} assigns exactly $k$ vertices to each $I_i$.
\end{cor}
\begin{proof}
Fix $i \in [n]$. It follows from the proof of Lemma~\ref{lem no fourth vertex} that in applying Algorithm~\ref{alg:put GN on WLD} to construct the set $I_i$, the algorithm can never reach the fourth vertex of a propagator's support (ordered with respect to $<_i$) without having already assigned $p$. Therefore if the algorithm starts at vertex $i$, all propagators must have been eliminated by the time it gets to vertex $i-1$, ensuring that the algorithm terminates and that $I_i$ contains exactly $k$ distinct elements.
\end{proof}

@@

Specifically, for each $i$ in the support of $p$, define the set
\[J_p^{W}(i) = \{m \in [n] \ | \ I^{W}_m(p) = i \},\]
the set of starting vertices $m$ such that $I_m^{W}$ assigns $p$ to $i$.

@@

Let $W = (\cP,[n])$ be an admissible Wilson loop diagram, and $p = (i,j) \in \cP$.  Then $J_p^{W}(i)$, $J_p^{W}(i+1)$, $J_p^{W}(j)$, and $J_p^{W}(j+1)$ are each non-empty cyclic intervals which partition $[n]$ and occur in the given cyclic order.
\end{lem}

@@

Consider the fourth case of Figure~\ref{fig 3 cases} first, as it is the easiest.  In this case, the vertices $\{l,l+1,l+2,l+3 \}$ which support $p$ in $W$ are non-supporting in $V$, so for every propagator $r$ in $V$ (including $q$) we have $J_r^{V} = J_r^{W}$; these are non-empty cyclic intervals in the correct order by the induction hypothesis.  Additionally, it is clear from Figure~\ref{fig 3 cases} that $J_p^{W}(l+a) = \{l+a\}$ for $a\in\{1,2,3\}$ and $J_p^{W}(l) = [n] \backslash \{l+1, l+2, l+3\}$. The result therefore holds in this case.

Now we proceed to consider the first three cases of Figure~\ref{fig 3 cases}.
We first describe $J_q^{V}(*)$, which can be handled identically for all three cases.

In $V$ there are no propagators inside $q$, so we see from Figure~\ref{fig V diagram} that
\[l, l+1, \ldots ,j \in J^{V}_q(j) \text{ (if $l$ exists) and }  j+1 \in J^{V}_q(j+1).\]
Note that $j+2 \not\in J^{V}_q(j+1)$ by Lemma~\ref{lem no fourth vertex}, so by the induction hypothesis we must have $J^{V}_q(j+1) = \{j+1\}$ and $j+2 \in J^{V}_q(i)$. Thus there exist vertices $d,e \in [j+2,i+1]$ with $d <e$, such that
\[J^{V}_q(i) = [j+2,d-1], \quad J^{V}_q(i+1) = [d,e-1], \quad J^{V}_q(j) = [e,j], \quad J^{V}_q(j+1) = \{j+1\},\]
and all intervals are non-empty.

We now consider what happens as we move from $V$ to each of the three remaining cases for $W$. We need to consider both $J^{W}_p$ and $J^{W}_r$ for $r \neq p$, since the addition of $p$ can have a knock-on effect on later steps in the algorithm.

@@

We first calculate $I^{W}_w$ for a starting vertex ${w \in [n]\backslash \{l, l+1, \ldots,j,j+1\}}$. Note that $p$ has no effect on other propagators for starting vertices in this range, while the value of $I^{W}_w(p)$ depends on how soon $q$ is assigned to a vertex, i.e. on the value of $I^{V}_w(q)$. Thus, if $w\in J_q^{V}(i)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        \max\{i+1, m\} & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
    \]
    while if $w\in J_q^{V}(i+1)$ or $w\in J_q^{V}(j)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        l & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
      \]
We also need to understand $I^{W}_w$ for $w \in \{l,l+1,\ldots,j,j+1\}$. For the majority of these vertices, we use the following observation: if $p$ is the first propagator to be assigned a value by $I_w^{W}$, then the remainder of $I_w^{W}$ proceeds identically to the assignments of $I^{V}_{w+1}$.  Thus we have for any $0\leq b <a$
    \[
    I_{l+b}^{W}(r) = \begin{cases}
      l+b & \text{if } r=p\\
      I_{l+b+1}^{V}(r) & \text{if } r\neq p
    \end{cases}
    \]
Similarly, if $j$ is in the support of $p$, then we have
     \[
       I_j^{W}(r) = \begin{cases}
         j & \text{if $r=p$ and $j$ is in the support of $p$}\\
         I_{j+1}^{V}(r) & \text{if $r\neq p$ and $j$ is in the support of $p$}\\
       \end{cases}
       \]
If $j$ is not in the support of $p$, then we must be in the second case of Figure~\ref{fig 3 cases} with two vertices in the right hand region.  In this case, if we start the algorithm at $j$ we need to know whether there will be any unassigned propagators other than $p$ when we reach vertex $i$, so as to know what $p$ contributes.

Consider the WLD $X$ formed from $V$ by moving the second end of $q$ to the edge $j-1$ instead of $j$.  $X$ is still admissible since we have not decreased the support of any set of propagators, so the induction hypothesis applies to it as well.  Note that $J_{j}^{V}(r) = J_{j}^{X}(r)$ for all $r\neq p$ and $J_{j}^{X}(r) = J_{j+1}^{X}(r)$ for $r\neq q,p$.  Additionally $I_{j+1}^{X}(q) = i$ by the induction hypothesis applied to $X$, and so if we start at $j+1$ and assign propagators to vertices according to the algorithm, when we reach vertex $i$ in $X$ all propagators other than $q$ must have been assigned.  Therefore if we start at $j$ in $W$, we first assign $q$ to $j$ then proceed to assign as in $X$ starting at $j+1$, and hence when we get to $i$ the only remaining unassigned propagator is $p$. Therefore
       \[
       I_j^{W}(r) = \begin{cases}

       m & \text{if $r =p$ and $j$ is not in the support of $p$} \\
       I_{j}^{V}(r) & \text{if $r\neq p$ and $j$ is not in the support of $p$}\end{cases}
       \] 
Finally, we consider what happens when we start the algorithm at vertex $j+1$. If $j+1$ is in the support of $p$ then we can argue as above to get
\[
       I_{j+1}^{W}(r)  = \begin{cases}
         j+1 & \text{if $r=p$ and $j+1$ is in the support of $p$}\\
         I_{j+2}^{V}(r) & \text{if $r\neq p$ and $j+1$ is in the support of $p$}.
       \end{cases}
       \]
Now suppose $j+1$ is not in the support of $p$. If we start at $j+1$ we need to know whether there are any unassigned propagators supported on edge $i$ when we reach vertex $i$. We already know that $J_q^{V}(j+1) = \{j+1\}$; in particular this means that $q$ contributes $i$ in $I_{j+2}^{V}$.  However the construction of $I^{V}_{j+1}$ first associates $q$ to $j+1$ and then proceeds identically to $I^{V}_{j+2}$.  In particular if $i$ was assigned in $I^{V}_{j+1}$, then it would not be available to assign to $q$ in $I^{V}_{j+2}$ as all other propagators supported at $i$ in $V$ come before $q$.
